### 2019-11-26 Mathematical Statistics/ms_ch_03.rmd
mathematical statistics, ch3 work
mathematical statistics, ch3 work

ua.delay <- Flights[Carrier == "UA"]$Delay
aa.delay <- Flights[Carrier == "AA"]$Delay

observed.mean <- mean(ua.delay)
observed.sum <- sum(ua.delay)
observed.diff <- mean(ua.delay) - mean(aa.delay)

@@

for(i in 1:N)
{
   index <- sample(nrow(Flights), length(ua.delay), replace = F)

@@

>>>>

UA.Delay <- Flights[Carrier == "UA"]$Delay
AA.Delay <- Flights[Carrier == "AA"]$Delay

observed.mean <- mean(UA.Delay)
observed.sum <- sum(UA.Delay)
observed.diff <- mean(UA.Delay) - mean(AA.Delay)

@@

for(i in 1:N)
{
   index <- sample(nrow(Flights), length(UA.Delay), replace = F)

@@

### 3.8

In the flight delays case study in Section 1.1,

a.) Find the trimmed mean of the delay times for United Airlines and American Airlines.

trim.amount <- .25
UA.trimmed <- mean(UA.Delay, trim = trim.amount)
AA.trimmed <- mean(AA.Delay, trim = trim.amount)

observed <- UA.trimmed - AA.trimmed

pretty_kable(data.table( UA = UA.trimmed, AA = AA.trimmed), "Trimmed Means")

b.) Conduct a two-sided test to see if the difference in trimmed means is statistically significant.

```{r, echo = T, fig.height=3.5, fig.width=8}
N <- 10e2 - 1
results <- numeric(N)

for(i in 1:N)
{
   index <- sample(nrow(Flights), nrow(Flights[Carrier == "UA"]), replace = F)
   results[i] <- mean(Flights[index]$Delay, trim = trim.amount) - mean(Flights[-index]$Delay, trim = trim.amount)
}

ggplot(data.table(results)) +
   geom_histogram(aes(results, fill = ..count..), bins = 30) +
   geom_vline(xintercept = observed, col = "darkorange", linetype = 3, lwd = 1.2) +
   scale_y_continuous(labels = comma) +
   labs(title = paste0("Flight Delay Time Trimmed Mean vs Observed, p=", round(p, 5)), 
        subtitle = paste0("Observed Value: ", round(observed, 4)))
```

### 3.9

In the flight delays case study in Section 1.1,

a.) Compute the proportion of times the flights in May and in June were delayed more than 20 min.

delay20_month <- Flights[, .(Delay20 = sum(Delay > 20) / .N), by = Month]

pretty_kable(delay20_month, "Delayed by more than 20 min, by month")

Conduct a two-sided test for statistical significance.

observed <- delay20_month[Month == "May"]$Delay20 - delay20_month[Month == "June"]$Delay20

for( i in 1:N)
{
   index <- sample(nrow(Flights), nrow(Flights[Month == "May"]), replace = F)
   results[i] <- as.numeric( Flights[index, .(Delay = sum(Delay > 20) / .N)] - Flights[-index, .(Delay = sum(Delay > 20) / .N)] )
}

P-value: `r round(p, 4)`, which is statistically significant.

b.) Compute the ratio of the variances in the flight delay times in May and in June.

observed <- var(Flights[Month == "May"]$Delay) - var(Flights[Month == "June"]$Delay)

for( i in 1:N)
{
   index <- sample(nrow(Flights), nrow(Flights[Month == "May"]), replace = F)
   results[i] <- var( Flights[index, .(Delay)] ) - var( Flights[-index, .(Delay)] )
}

Is this evidence that the true ratio is not equal to 1, or could this be due to chance variability?

The variance appear to be due to random chance, so there does appear to be a stastical significance between the two months.

Conduct a two-sided test to check.

P-value: __`r p`__

### 3.10

In the black spruce case study in Section 1.10, seedlings were planted in plots that were either subject to competition (from other plants), or not.

Use the data set _Spruce_ to conduct a test to see if the mean difference is how much the seedlings grow (in height) over the corse of the study under these two treatments is stastically significant.

```{r, echo = T}
Spruce <- data.table(read.csv(paste0(data.dir, "Spruce.csv"),
                               header = T))

observed <- mean( Spruce[Competition == "NC"]$Ht.change ) - mean( Spruce[Competition == "C"]$Ht.change )

for(i in 1:N)
{
   index <- sample(nrow(Spruce), nrow(Spruce[Competition == "NC"]), replace = F)
   resutls <- mean( Spruce[index]$Ht.change ) - mean( Spruce[-index]$Ht.change )
}

p <- min(1, 2 * (sum(results[results >= observed]) + 1) / ( N + 1))

There is stastical significance in between the heights of the two groups (Competition / No-competition).

P-value: __`r p`__

### 3.11

The file _Phillies2009_ contains data from the 2009 season for the baseball team the Philadelphia Phillies.

```{r, echo = T}
Phillies <- data.table(read.csv(paste0(data.dir, "Phillies2009.csv"),
                               header = T))
```

a.) Compare the empirical distribution functions of the number of strike-outs per game (_StrikeOuts_) for games played at home and games played away (_Location_).

```{r, echo = T, fig.width=8, fig.height=3.5}

ggplot(Phillies, aes(StrikeOuts, color = Location)) +
   stat_ecdf(stat = "point") +
   labs(title = "Strike-outs / Game, by Location", y = "Percent")

b.) Find the mean number of strike-outs per game for the home and the away games.

```{r, echo = T}
strikeouts <- Phillies[, .(StrikeOuts = mean(StrikeOuts)), by = Location]

pretty_kable(strikeouts, "StrikeOuts by Location")

observed <- mean(strikeouts[Location == "Away"]$StrikeOuts) - mean(strikeouts[Location == "Home"]$StrikeOuts)

```
c.) Perform a permutation tests to see if the differences in means is statistically significant.

for( i in 1:N)
{
   index <- sample(nrow(Phillies), nrow(Phillies[Location == "Home"]), replace = F)
   results[i] <- mean(Phillies[index]$StrikeOuts) - mean(Phillies[-index]$StrikeOuts)
}

p <- min(1, ( sum(results >= observed) + 1) / ( N + 1) )

There does not appear to be a statistically significat relationship between the number of home and away strikeouts.

P-value: __`r p`__
